Brownian dynamics simulation of cytoplasmic dynein's powerstroke reproduces bidirectional stepping

### 2018-11-07 papers/paper/paper.tex

@@

\textbf{Processive, forward-directed motility}\\
The model was time-evolved using Euler's method. When simulated, the model produces processive, forward-directed motion. Figure \ref{fig:trajectory} shows a short time window of the model's dynamics. The model is capable of taking both forward and backwards steps, but is biased in the forward direction.\\

\textbf{Stepping dynamics}
The stepping dynamics of the model reproduce experimental results. The model is capable of producing a wide variety of step lengths, from 48nm forward to 40nm backwards, as in Figure \ref{fig:behavior}.a. The onebound and bothbound times correspond well to the theoretical values, as in \ref{fig:behavior}.b and c. The tension-gating mechanism is capable of producing a lagging-head favor, as shown in \ref{fig:behavior}.d.

%% -Briefly how we simulated the model
%% -Model is forwards-directed and processive (trajectory plot)
%% -Tuning the parameters to Yildiz data (kb/kub from theoretical OB/BB time and step length histograms for cm/cb/ct and leading/lagging fractions for c)
%% -Does the tuned model walk similarly to Yildiz dynein? (also the lagging vs step length)

@@

@@

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=\linewidth]{../../plots/paper_force_vs_velocity.pdf}
\caption{\textbf{Impact of force on model velocity.} Model response to force in the $+\hat{x}$ direction applied to tail domain. Velocity calculated using displacement after 2 seconds of simulation at standard parameters. Behavior compared with a similar experiment on a cytoplasmic dynein construct using optical tweezers from Gennerich \textit{et. al.} \cite{responsetoload}.}
\label{fig:}
\end{figure}

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=\linewidth]{../../plots/paper_initial_vs_final_displacement.pdf}
  \caption{\textbf{Analysis of stepping independence.} Heatmap of MTBD displacement before and after a stepping cycle. Displacement defined as the unbinding MTBD's x coordinate minus the stationary MTBD's x coordinate, calculated before and after the step. Structure in the heat map suggests steps are not independent, but rather that correlation exists between starting position and step size.}
\label{fig:}
\end{figure}

>>>>

Motion equations were derived for the pre-powerstroke and post-powerstroke states of the model. The model was simulated by starting in the equilibrium post-powerstroke position and using Euler's method to time-evolve the model, using a timestep of 1e-10 \textit{s}. The model was transitioned between post-powerstroke and pre-powerstroke states using unbinding and binding rates. Model parameters were estimated from various experimental results. Spring constants were chosen to fit the model stepping size to experiment, as outlined in Figure S\ref{fig:supp-model-fit}. Model rate constants were fit to theoretical calculations outlined in the Methods section. The tension-gating parameter c was fit as shown in Figure \ref{fig:tensiongating}. 32 5-second simulations were run, each initialized with a different random seed. Each simulation was forward directed, took both forward and backward steps, and displayed stepping dynamics similar to that seen in experiment.\\

  As seen in Figure \ref{fig:trajectory}, the model is highly variant in trajectory, but biased to moving in the $+\hat{x}$ direction, corresponding to the microtubule minus-end. When unbound the free MTBD diffuses both parallel and perpendicular to the microtubule, reaching up to 50 \textit{nm} above the track. Though the unbound MTBD freely approaches the microtubule, it does not always bind, as seen in the $7-9 \mu s$ part of the simulation. The cartoons in Figure \ref{fig:trajectory} reveal the model taking on many conformations, revealing the model's structural flexibility. The model seems to prefer alternating hand-over-hand stepping, but occasionally takes steps in a semi-inchworm fashion, as seen at time $5.3 \mu s$ in Figure \ref{fig:trajectory}.\\

@@

The model was then compared to TIRF studies of dynein stepping in DeWitt \textit{et. al.} \cite{yildizpaper}. Figure \ref{fig:behavior}.a shows 8 representative model trajectories compared to experimental dynein. The eight simulations track experimental very well over 5 seconds, and average to the same position. The eight simulations end up distributed over a 200 \textit{nm} area, showing the variation in the model stepping extends to a variation in velocity in the long term. The step size distribution for the 32 simulations are shown in Figure \ref{fig:behavior}.b). The model is capable of taking both positive and negative steps, with steps extending from -40 \textit{nm} to 50 \textit{nm}, matching experimental dynein. The most probable step is 12 \textit{nm}, with an average step size of 6.2 \textit{nm}. This matches the peak experimental step size window of 7-13 \textit{nm}. The major deviation in model and experiment is the density of model steps around 0nm, which is possibly caused by an inability of the step detection algorithm to detect small steps. The pre-powerstroke and post-powerstroke state occupation times are shown in Figure \ref{fig:behavior}.c and d.). The model prefers long pre-powerstroke times, but is capable of taking very fast steps. The post-powerstroke time has a heavy-tailed distribution. Experimental technique does not resolve time spent in pre- and post-powerstroke, but we outline a technique for calculating these values for experimental trajectories in Methods. These theoretical values correspond well to model behavior.\\

@@

Dependence of the unbinding MTBD on its relation to its partner MTBD is studied in Figure \ref{fig:tensiongating}. Figure \ref{fig:tensiongating}.a shows a heat map of initial displacement, defined as how much the stepping MTBD is ahead of the stationary MTBD, vs the step size. A line of best fit is constructed with intercept $b=5.7$ and slope $m=-1$. This indicates a negative trend where large positive displacements are likely to cause backwards steps, and negative displacements are likely to cause forward steps. This is consistent with the idea of tension gating, where tension between the two motor heads drives unbinding to relieve tension. The negative trend is reproduced in DeWitt \textit{et. al}, which shows a slope of $m=-0.4$. Figure \ref{fig:tensiongating}.b shows the unbinding probability of the lagging head with distance to the leading head. As the distance becomes greater the lagging head is more and more likely to unbind, again consistent with tension gating. This trend also very closely tracks DeWitt \textit{et. al}.\\

@@

The effect of force on the model was then assessed by applying an artificial force on the tail domain of the model in the $+\hat{x}$ direction. The results of this study are shown in Figure \ref{fig:force}. An expected decrease in velocity is seen as stronger and stronger negative forces are applied to the model, pulling it against its natural walking direction. Unlike experiment, a reversal in velocity is not seen for this trial, indicating that perhaps the dynamics of force-mediated velocity reversal are more complicated than what this simplified model can handle. The results are consistent with an experiment done by Grotjahn \textit{et. al} where force is applied via optical tweezers to dynein constructs and their velocity is assayed. It will need to be studied why odd behavior is seen in the mdoel for positive forces.\\

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=\linewidth]{../../plots/paper_force_vs_velocity.pdf}
\caption{\textbf{Impact of force on model velocity.} Model response to force in the $+\hat{x}$ direction applied to tail domain. Velocity calculated using displacement after 2 seconds of simulation at standard parameters. Behavior compared with a similar experiment on a cytoplasmic dynein construct using optical tweezers from Gennerich \textit{et. al.} \cite{responsetoload}.}
\label{fig:force}
\end{figure}

The independence of dynein steps is then studied in Figure \ref{fig:independence}. Whether dynein works like a Markovian process, where each step is a history-free, is an interesting question. Figure \ref{fig:independence} demonstrates the impact of initial displacement of MTBDs to their final displacement, before and after a step is taken. The asymmetry in the plot is indicative that the initial configuration of the motor does have significant impact on the later state. The Brownian forces act to mitigate momentum and prevent information transfer through motion. However, the binding orientation of the dynein model to the microtubule is stable through time, and can influence the binding orientation of dynein to the microtubule at a later time. There seems to be a roughly tension-gated process occurring where negative displacements initially lead to positive displacements later.\\

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=\linewidth]{../../plots/paper_initial_vs_final_displacement.pdf}
  \caption{\textbf{Analysis of stepping independence.} Heatmap of MTBD displacement before and after a stepping cycle. Displacement defined as the unbinding MTBD's x coordinate minus the stationary MTBD's x coordinate, calculated before and after the step. Structure in the heat map suggests steps are not independent, but rather that correlation exists between starting position and step size.}
\label{fig:independence}
\end{figure}
